---
title: "Chapter 4"
output: html_document
date: "2022-11-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = '/Users/ryanfield/Downloads/Paradigm')
##################################################################
##                        Load Libraries:                       ##
##                     rio: for data import                     ##
##           tidyverse: for data cleaning / wrangling           ##A
##               lubridate: for date manipulation               ##
##                     JM: for joint models                     ##
##           conflicted: to resolve package conflicts           ##
##################################################################

#############################
library(rio)
library(tidyverse)
library(lubridate)
library(conflicted)
library(mice)
library(survAUC)

########################################################################
##  Prefer functions from dplyr over JM for easier data manipulation  ##
########################################################################
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")

```


```{r import-data}
#########################################################################################
##  Import Biomarkers from abio.dta dataset and filter natriuretic peptides (NTProBNP) ##
#########################################################################################

NatriureticPeptides <- import("STATA/CLCZ696B_CLCZ696B2314_glasgow_analysis_datasets_M8_25Sep2014/abio.dta")

unique(NatriureticPeptides$PARNAM1C)

ntprobnp <- NatriureticPeptides %>% filter(PARNAM1C == "NTPBNP")

NatriureticPeptides <- NULL

ntprobnp <- ntprobnp %>% select(sid1a = SID1A,
                                ntprobnp_value = LABRSL1N,
                                sample_date = SMPCOL1D,
                                visit = VISNAM1A)

```

```{r import-baseline}
##################################################################
##      Import baseline variables form from master dataset      ##
##################################################################

baseline <- import("Master/paradigmhf master version 11 compatible.dta")

##################################################################
##             Select Variables to keep in baseline             ##
##     where possible use variables from randomisation visit     ##
##################################################################

baseline <- baseline %>% select(sid1a,
                                randate,
                                t2dth,
                                t2cnpt,
                                t2hhf,
                                c4dth,
                                c4cnpt,
                                c4hhf,
                                age = age_1n,
                                region,
                                sex = sex1c,
                                trt = trtreg1c,
                                bmi = bmi_b2,
                                NYHA = nyh_b2,
                                eGFR = gfr_b2,
                                ejf = ejf_b1,
                                diabetes = dbt_flg2,
                                atrial_fibrillation = af_flg1,
                                prior_hf = phh_flg1,
                                prior_mi = mi_flg1,
                                prior_stroke = stk_flg1,
                                ischemic_hf = ich_flg1,
                                sbp = sbp_b2,
                                heart_rate = hr_b2
)

```

# Baseline Minipulation
```{r filter-baseline}

# View Missing Data
md.pattern(baseline)

# Recode Region to factor
baseline <- baseline %>% mutate(region = as.factor(case_when(
  region == 1 ~ "North American",
  region == 2 ~ "Latin America",
  region == 3 ~ "Western Europe",
  region == 4 ~ "Central Europe",
  region == 5 ~ "Pacific Asia/Pacific and Other"
)))

# Recode Treatment
baseline <- baseline %>% mutate(trt_drug = as.factor(case_when(
  trt == "A" ~ "LCZ",
  trt == "B" ~ "Enalapril"
)))

# Recode NYHA Class to factor
baseline <- baseline %>% mutate(NYHA_class = as.factor(case_when(
  NYHA == "1" ~ "CLASS_1",
  NYHA == "2" ~ "CLASS_2",
  NYHA == "3" ~ "CLASS_3",
  NYHA == "4" ~ "CLASS_4"
)))

# Recode Sex as Factor
baseline <- baseline %>% mutate(sex = as.factor(
  if_else(sex == 1, "MALE", "FEMALE")))

# Recode diabetes as Factor
baseline <- baseline %>% mutate(diabetes = as.factor(
  if_else(diabetes == 1, "TRUE", "FALSE")))

# Recode AF as Factor
baseline <- baseline %>% mutate(atrial_fibrillation = as.factor(
  if_else(atrial_fibrillation == 1, "TRUE", "FALSE")))

# Recode Prior Hospitalisation for Heart Failure as Factor
baseline <- baseline %>% mutate(prior_hf = as.factor(
  if_else(prior_hf == 1, "TRUE", "FALSE")))

# Recode Prior MI as Factor
baseline <- baseline %>% mutate(prior_mi = as.factor(
  if_else(prior_mi == 1, "TRUE", "FALSE")))

# Recode Prior Stroke as Factor
baseline <- baseline %>% mutate(prior_stroke = as.factor(
  if_else(prior_stroke == 1, "TRUE", "FALSE")))

# Recode Time to Composite Event to 28 Calendar Day Months
baseline <- baseline %>% mutate(composite_month = t2cnpt / 28)

# Invert Censor For Composite Event to be 1 for Event
baseline <- baseline %>% mutate(composite_status = if_else(c4cnpt == 1, 0, 1))

# Recode Ischemic HF to Factor
baseline <- baseline %>% mutate(ischemic_hf = as.factor(if_else(ischemic_hf == 1, "YES", "NO")))

# Remove Missing for Ischemic HF
baseline <- baseline %>% filter(!is.na(ischemic_hf))
```

```{r filter-nt-probnp}

# Join with required baseline characteristics
ntprobnp <- ntprobnp %>% left_join(select(baseline, sid1a, trt_drug, eGFR, randate, age, sex, bmi, atrial_fibrillation, ischemic_hf), by="sid1a")

#################################################################
##                        Log NTProBNP                         ##
##  and filter any patients who did not undergo randomisation  ##
#################################################################
ntprobnp <- ntprobnp %>% mutate(logNTProBNP = log(ntprobnp_value)) %>%
  filter(!is.na(randate)) 

#######################################################################
##  Calculate Months Since Randomisation using 28 day calendar month  ##
#######################################################################
ntprobnp <- ntprobnp %>% 
  mutate(sample_date_formatted = as_date(sample_date, format = "%d%b%Y")) %>%
  mutate(sample_days = interval(randate, sample_date_formatted) %/% days(1)) 

#############################################################################################
##  Use Screening NT-ProBNP for Baseline, for patients without NT-ProBNP at Randomisation  ##
#############################################################################################

ntprobnp <- ntprobnp %>% arrange(sid1a) %>% group_by(sid1a) %>% arrange(sample_days, .by_group = TRUE)

ntprobnp <- ntprobnp %>% mutate(n_obs = n(),
                                last_visit = last(sample_days))

ntprobnp <- ntprobnp %>% mutate(sample_days = if_else(visit == "Screening" & last_visit < 0, 0, sample_days),
                                from_screening = if_else(visit == "Screening" & last_visit < 0, TRUE, FALSE))

ntprobnp <- ntprobnp %>% mutate(sample_month = sample_days / 28)

ntprobnp <- ntprobnp %>% filter(sample_month >= 0)

ntprobnp <- ntprobnp %>% filter(!is.na(eGFR))

baseline <- baseline %>% filter(sid1a %in% ntprobnp$sid1a)

```

```{r remove-missing-values}

# View Missing Data
missing_data <- ntprobnp[!complete.cases(ntprobnp),]
missing_data_baseline <- baseline[!complete.cases(baseline),]

# Filter patients missing NYHA Class, BMI, eGFR or baseline NT-ProBNP
baseline <- baseline %>% filter(!is.na(NYHA_class)) %>%
  filter(!is.na(bmi)) %>% filter(!is.na(ejf))

# View Missing Data
missing_data_baseline <- baseline[!complete.cases(baseline),]

# View Missing Data
missing_data <- ntprobnp[!complete.cases(ntprobnp),]

# Reorder Rows to match baseline with repeated measures of NT-ProBNP
ntprobnp <- ntprobnp[order(ntprobnp$sid1a),]
row.names(ntprobnp) <- 1:nrow(ntprobnp)
baseline <- baseline[order(baseline$sid1a),]
row.names(baseline) <- 1:nrow(baseline)

# Filter patients from each dataset who have been removed from either
ntprobnp <- ntprobnp %>% filter(sid1a %in% baseline$sid1a)
baseline <- baseline %>% filter(sid1a %in% ntprobnp$sid1a)

# Export the baseline dataset
export(baseline, "r-scripts-ryan/Final-Analysis/ALL-NT-PROBNP/Data/baseline.Rds")
```

```{r survival-data-set}
# Make a new dataset of repeat measurements with all baseline characteristics
ntprobnp_survival <- ntprobnp %>% left_join(baseline, by = c("sid1a", "trt_drug", "eGFR", "randate", "age", "sex", "bmi", "atrial_fibrillation", "ischemic_hf")) %>% group_by(sid1a)

# Make a baseline for this new dataset by removing repeated measurements
ntprobnp_baseline <- ntprobnp_survival %>% group_by(sid1a) %>% arrange(sample_month, by_group = TRUE)
ntprobnp_baseline <- ntprobnp_baseline[!duplicated(ntprobnp_baseline$sid1a),]

# Export this new dataset
export(ntprobnp_survival, "r-scripts-ryan/Final-Analysis/ALL-NT-PROBNP/Data/ntprobnp_survival.Rds")

```

```{r spaghetti-plots, fig.dim = c(20, 20), warning=FALSE}
# Set seed
set.seed(1234, "Mersenne-Twister", sample.kind = "Rejection")

# Add Number of Repeated Measurements of NT-ProBNP
ntprobnp <- ntprobnp %>% group_by(sid1a) %>% arrange(sample_days, .by_group = TRUE) %>% mutate(n_obs = n())
# Export NT-ProBNP Dataset
export(ntprobnp, "r-scripts-ryan/Final-Analysis/ALL-NT-PROBNP/Data/ntprobnp.Rds")

##################################################################
##                   Stratified Randomisation                   ##
##                         and sampling                         ##
##################################################################
ntprobnp_1 <- ntprobnp %>% filter(n_obs == 1)
ntprobnp_2 <- ntprobnp %>% filter(n_obs == 2)
ntprobnp_3 <- ntprobnp %>% filter(n_obs >= 3)

u_ids_1 <- ntprobnp_1[!duplicated(ntprobnp_1$sid1a),]
u_ids_2 <- ntprobnp_2[!duplicated(ntprobnp_2$sid1a),]
u_ids_3 <- ntprobnp_3[!duplicated(ntprobnp_3$sid1a),]

u_ids_1.Enalapril <- u_ids_1 %>% filter(trt_drug == "Enalapril")
u_ids_2.Enalapril <- u_ids_2 %>% filter(trt_drug == "Enalapril")
u_ids_3.Enalapril <- u_ids_3 %>% filter(trt_drug == "Enalapril")

u_ids_1.LCZ <- u_ids_1 %>% filter(trt_drug == "LCZ")
u_ids_2.LCZ <- u_ids_2 %>% filter(trt_drug == "LCZ")
u_ids_3.LCZ <- u_ids_3 %>% filter(trt_drug == "LCZ")

u_ids_sample_1.Enalapril <- sample(u_ids_1.Enalapril$sid1a, 3)
u_ids_sample_2.Enalapril <- sample(u_ids_2.Enalapril$sid1a, 6)
u_ids_sample_3.Enalapril <- sample(u_ids_3.Enalapril$sid1a, 12)

u_ids_sample_1.LCZ <- sample(u_ids_1.LCZ$sid1a, 3)
u_ids_sample_2.LCZ <- sample(u_ids_2.LCZ$sid1a, 6)
u_ids_sample_3.LCZ <- sample(u_ids_3.LCZ$sid1a, 12)

u_ids_sample <- c(u_ids_sample_1.Enalapril, u_ids_sample_2.Enalapril, u_ids_sample_3.Enalapril,
                  u_ids_sample_1.LCZ, u_ids_sample_2.LCZ, u_ids_sample_3.LCZ)

ntprobnp_sample <- ntprobnp %>% filter(sid1a %in% u_ids_sample) %>% group_by(sid1a)

# Plot the randomly stratified patients
splot <- ggplot(data = ntprobnp_sample, aes(x = sample_month, y = logNTProBNP, color = trt_drug)) + geom_point() + geom_line() + facet_wrap(~sid1a) +
  xlab("(Log) NTProBNP") +
  ylab("Time (Months)") +
  ggtitle("Longitudinal Profiles of (Log) NTProBNP for 42 Sampled Patients") + 
  labs(colour = "Treatment Group")

splot

```

# Baseline Characteristics Table
```{r baseline-stats-q1-q3}

library(table1)

table1(~
  age +
  sex +
  region +
  bmi + 
  eGFR + 
  ejf +
  NYHA_class +
  diabetes + 
  sbp +
  heart_rate +
  atrial_fibrillation +
  ischemic_hf +
  prior_hf +
  prior_mi + 
  prior_stroke | trt_drug,
  data = baseline,
  render.continuous=c(.="Median [Q1, Q3]" ))

```

# NT-ProBNP at Baseline
```{r nt-probnp-0-q1-q3}
ntprobnp_at_baseline <- ntprobnp %>% group_by(sid1a) %>% arrange(sid1a, sample_month, .by_group = T) %>% filter(row_number()==1)

table1(~ exp(logNTProBNP) | trt_drug,
  data = ntprobnp_at_baseline,
  render.continuous=c(.="Median [Q1, Q3]" ))

```

# Fit LME
```{r lme}
library(JMbayes2)

#################################################################
##                           Fit LME                           ##
#################################################################
sample_month_quantiles <- quantile(ntprobnp_1$sample_month)
sample_month_quantiles


lme.3 <- lme(logNTProBNP ~ 
               ns(sample_month, df = 3, Boundary.knots = c(1, 17.25)) * trt_drug +
                age +
                atrial_fibrillation +
                bmi,
           random = ~ ns(sample_month, df = 3, Boundary.knots = c(1, 17.25)) | sid1a,
           data = ntprobnp,
           control = lmeControl(opt = 'optim'), method = "ML")

export(lme.3, "r-scripts-ryan/Final-Analysis/ALL-NT-PROBNP/Data/lme.3.Rds")

```

# Fit Cox PH
```{r, coxphcomp}
coxPH.comp.fit <- coxph(Surv(composite_month, composite_status) ~
                     trt_drug +
                     age +
                     sex +
                     region +
                     bmi + 
                     eGFR + 
                     ejf +
                     NYHA_class +
                     diabetes + 
                     sbp +
                     heart_rate +
                     atrial_fibrillation +
                     ischemic_hf +
                     prior_hf +
                     prior_mi + 
                     prior_stroke,
                   data = baseline,
                   x = TRUE)

export(coxPH.comp.fit, "r-scripts-ryan/Final-Analysis/ALL-NT-PROBNP/Data/coxPH.comp.fit.Rds")

```

# Fit Joint Models
```{r, jmcomp}

##################################################################
##                      Fit Joint Models                        ##
##               Joint Model 1 =  jointFit.comp.1               ##
##               Joint Model 2 =  jointFit.comp.4               ##
##               Joint Model 3 =  jointFit.comp.6               ##
##################################################################

jointFit.comp.1 <- jm(coxPH.comp.fit, lme.3, time_var = "sample_month")

summary(jointFit.comp.1)

jointFit.comp.4 <- jm(coxPH.comp.fit, lme.3, time_var = "sample_month",
                 functional_forms = ~ value(logNTProBNP) + slope(logNTProBNP))
summary(jointFit.comp.4)

jointFit.comp.6 <- jm(coxPH.comp.fit, lme.3, time_var = "sample_month",
                 functional_forms = ~ area(logNTProBNP))
summary(jointFit.comp.6)

export(jointFit.comp.1, "r-scripts-ryan/Final-Analysis/ALL-NT-PROBNP/Data/jointFit.comp.1.Rds")
export(jointFit.comp.4, "r-scripts-ryan/Final-Analysis/ALL-NT-PROBNP/Data/jointFit.comp.4.Rds")
export(jointFit.comp.6, "r-scripts-ryan/Final-Analysis/ALL-NT-PROBNP/Data/jointFit.comp.6.Rds")

```

# Setup Data for Avarage Trajectories
```{r setup-average-trajectories}

avg_logNTProBNP.lcz <- ntprobnp %>% filter(trt_drug == 'LCZ')
avg_logNTProBNP.lcz <- median(avg_logNTProBNP.lcz$logNTProBNP)

avg_logNTProBNP.enalapril <- ntprobnp %>% filter(trt_drug == 'Enalapril')
avg_logNTProBNP.enalapril <- ntprobnp %>% filter(sample_month == 0, trt_drug == 'Enalapril')
avg_logNTProBNP.enalapril <- median(avg_logNTProBNP.enalapril$logNTProBNP)

##################################################################
##         Function to return the most frequent Factors         ##
##################################################################
mostfrequentFactor <- function(column){
  mff <- data.frame(table(column))
  mff[which.max(mff$Freq),]$column
}

##################################################################
##   Make a Dataset for the Average Patient in each treatment   ##
##################################################################
avg_ntprobnp<- with(ntprobnp_baseline,
                              expand.grid(
                                sid1a = 0,
                                sample_month = 0,
                                trt_drug = factor("LCZ", levels = levels(ntprobnp_survival$trt_drug)),
                                eGFR = median(eGFR),
                                age = median(age),
                                sex = mostfrequentFactor(sex),
                                region = mostfrequentFactor(region),
                                bmi = median(bmi),
                                ejf = median(ejf),
                                sbp = median(sbp),
                                heart_rate = median(heart_rate),
                                NYHA_class = mostfrequentFactor(NYHA_class),
                                diabetes = mostfrequentFactor(diabetes),
                                atrial_fibrillation = mostfrequentFactor(atrial_fibrillation),
                                ischemic_hf = mostfrequentFactor(ischemic_hf),
                                prior_hf = mostfrequentFactor(prior_hf),
                                prior_mi  = mostfrequentFactor(prior_mi),
                                prior_stroke = mostfrequentFactor(prior_stroke),
                                logNTProBNP = 0
                              ))


avg_ntprobnp_2<- with(ntprobnp_baseline,
                              expand.grid(
                                sid1a = 1,
                                sample_month = 0,
                                eGFR = median(eGFR),
                                trt_drug = factor("Enalapril", levels = levels(ntprobnp_survival$trt_drug)),
                                age = median(age),
                                sex = mostfrequentFactor(sex),
                                region = mostfrequentFactor(region),
                                bmi = median(bmi),
                                ejf = median(ejf),
                                sbp = median(sbp),
                                heart_rate = median(heart_rate),
                                NYHA_class = mostfrequentFactor(NYHA_class),
                                diabetes = mostfrequentFactor(diabetes),
                                atrial_fibrillation = mostfrequentFactor(atrial_fibrillation),
                                ischemic_hf = mostfrequentFactor(ischemic_hf),
                                prior_hf = mostfrequentFactor(prior_hf),
                                prior_mi  = mostfrequentFactor(prior_mi),
                                prior_stroke = mostfrequentFactor(prior_stroke),
                                logNTProBNP = 0
                              ))


library(kableExtra)
kable(avg_ntprobnp)

```

# Plot Average Trajectories from Joint Model 1
```{r Average_Trajectories}

#################################################################
##          Predict Average Values for Each Treatment          ##
#################################################################
l_pred.lcz <- predict(jointFit.comp.1, newdata = avg_ntprobnp,
            times = seq(0, 17.25, length.out = 51),
            type = "mean_subject",
            return_newdata = T)

l_pred.enalapril <- predict(jointFit.comp.1, newdata = avg_ntprobnp_2,
            times = seq(0, 17.25, length.out = 51),
            type = "mean_subject",
            return_newdata = T)

# Extract the predictions
lcz <- l_pred.lcz$newdata2
enalapril <- l_pred.enalapril$newdata2

# Bind these two dataset together
lcz_enalapril <- rbind(lcz, enalapril)

# Plot the average Trajectories
ggplot(data = ntprobnp, mapping = aes(x = sample_month, y = logNTProBNP, color = trt_drug)) + 
         geom_point(mapping = aes(group = sid1a)) + geom_line(mapping = aes(group = sid1a), alpha = 0.5) +
  xlim(0,12) + ylim(6.25,8) + 
  geom_ribbon(aes(ymin=low_logNTProBNP, ymax=upp_logNTProBNP), linetype=2, alpha=0.5, data = lcz_enalapril, fill = 'white') + 
  geom_line(data = lcz_enalapril, mapping = aes(x = sample_month, y = pred_logNTProBNP, color = trt_drug), alpha = 1) + labs(color = "Treatment") + ggtitle("Average Trajectories of (Log) NTProBNP by Treatment Group") + xlab("Sample Time (Months)") + ylab("(Log) NT ProBNP")

```

# Predicted Average Trajectories from Joint Model 2
```{r Average_Trajectories-4}

#################################################################
##          Predict Average Values for Each Treatment          ##
#################################################################
l_pred.lcz <- predict(jointFit.comp.4, newdata = avg_ntprobnp,
            times = seq(0, 17.25, length.out = 51),
            type = "mean_subject",
            return_newdata = T)

l_pred.enalapril <- predict(jointFit.comp.4, newdata = avg_ntprobnp_2,
            times = seq(0, 17.25, length.out = 51),
            type = "mean_subject",
            return_newdata = T)

# Extract the predictions
lcz <- l_pred.lcz$newdata2
enalapril <- l_pred.enalapril$newdata2

# Bind these two datasets together
lcz_enalapril <- rbind(lcz, enalapril)

# Plot the average Trajectories
ggplot(data = ntprobnp, mapping = aes(x = sample_month, y = logNTProBNP, color = trt_drug)) + 
         geom_point(mapping = aes(group = sid1a)) + geom_line(mapping = aes(group = sid1a), alpha = 0.5) +
  xlim(0,12) + ylim(6.25,8) + 
  geom_ribbon(aes(ymin=low_logNTProBNP, ymax=upp_logNTProBNP), linetype=2, alpha=0.5, data = lcz_enalapril, fill = 'white') + 
  geom_line(data = lcz_enalapril, mapping = aes(x = sample_month, y = pred_logNTProBNP, color = trt_drug), alpha = 1) + labs(color = "Treatment") + ggtitle("Average Trajectories of (Log) NTProBNP by Treatment Group - Joint Model 2") + xlab("Sample Time (Months)") + ylab("(Log) NT ProBNP")

```

# Predicted Average Trajectories from Joint Model 3
```{r Average_Trajectories-6}

#################################################################
##          Predict Average Values for Each Treatment          ##
#################################################################
l_pred.lcz <- predict(jointFit.comp.6, newdata = avg_ntprobnp,
            times = seq(0, 17.25, length.out = 51),
            type = "mean_subject",
            return_newdata = T)

l_pred.enalapril <- predict(jointFit.comp.6, newdata = avg_ntprobnp_2,
            times = seq(0, 17.25, length.out = 51),
            type = "mean_subject",
            return_newdata = T)

# Extract the predictions
lcz <- l_pred.lcz$newdata2
enalapril <- l_pred.enalapril$newdata2

# Bind these two datasets together
lcz_enalapril <- rbind(lcz, enalapril)

# Plot the average Trajectories
ggplot(data = ntprobnp, mapping = aes(x = sample_month, y = logNTProBNP, color = trt_drug)) + 
         geom_point(mapping = aes(group = sid1a)) + geom_line(mapping = aes(group = sid1a), alpha = 0.5) +
  xlim(0,12) + ylim(6.25,8) + 
  geom_ribbon(aes(ymin=low_logNTProBNP, ymax=upp_logNTProBNP), linetype=2, alpha=0.5, data = lcz_enalapril, fill = 'white') + 
  geom_line(data = lcz_enalapril, mapping = aes(x = sample_month, y = pred_logNTProBNP, color = trt_drug), alpha = 1) + labs(color = "Treatment") + ggtitle("Average Trajectories of (Log) NTProBNP by Treatment Group - Joint Model 3") + xlab("Sample Time (Months)") + ylab("(Log) NT ProBNP")

```

# Produce longitudinal outcomes table
```{r longitudinal-outcomes}
library(kableExtra)

jm1.long <- as.data.frame(summary(jointFit.comp.1)$Outcome1) %>% rownames_to_column()
jm4.long <- as.data.frame(summary(jointFit.comp.4)$Outcome1) %>% rownames_to_column()
jm6.long <- as.data.frame(summary(jointFit.comp.6)$Outcome1) %>% rownames_to_column()

kable(jm1.long, caption = "JM1")
kable(jm4.long, caption = "JM4")
kable(jm6.long, caption = "JM6")

```

# Dynamic Predictions
```{r dynamic-predictions}
# Select Randomly sampled patient
ntprobnp_4 <- ntprobnp_survival %>% filter(sid1a == "3142_00004")

# Set the Survival Outcome to Null
ntprobnp_4$composite_month <- NULL
ntprobnp_4$composite_status <- NULL

# Function for dynamic prediction plots
dynamic_pred <- function(newData){
  for (i in seq_len(nrow(newData)))
  {
    # Predict Longitudinal Outcome
    l_pred.4 <- predict(jointFit.comp.4, newdata = newData[1:i,],
            times = seq(0, 24, length.out = 51),
            type = "subject_specific",
            return_newdata = T)
    # Predict Survival Outcome
    s_pred.4 <- predict(jointFit.comp.4, newdata = newData[1:i,],
              times = seq(0, 24, length.out = 51),
              process = "event",
              return_newdata = T)
    # Plot Longitudinal and Survival Outcomes
    plot(l_pred.4, s_pred.4,
       fun_long = exp,
           fun_event = function (x) 1 - x,
           ylab_event = "Survival Probability",
           ylab_long = "NT-ProBNP",
           xlab = "Follow-up Time (Months)",
           CI_long = T,
           CI_event = T,
           ylim_long_outcome_range = F)  + title(paste0("Observation ", i))
    # Print the Survival at 24 Months
    print(paste0('Survival at 24 Months: ', round(1 - s_pred.4[s_pred.4$sample_month == 24, ]$pred_CIF,2)))
    print(paste0('(',
                 round(1 - s_pred.4[s_pred.4$sample_month == 24, ]$low_CIF,2),
                 ' - ',
                 round(1 - s_pred.4[s_pred.4$sample_month == 24, ]$upp_CIF,2),
                 ')'
                 ))
    
  }
}

# Call the funciton to plot the dynamic predictions
dynamic_pred(ntprobnp_4)


```

# Joint Model 1 HR's
```{r jm1-hr}
summary.jm <- as.data.frame(summary(jointFit.comp.1)$Survival)
summary.jm <- summary.jm %>% mutate(HR = exp(Mean))
summary.jm <- round(summary.jm,3)
summary.jm <- summary.jm %>%mutate(P = ifelse(P == 0, '< 0.000', P))
summary.jm <- summary.jm %>% mutate(HR = paste0(round(exp(Mean), 2), '(', round(exp(`2.5%`),2), ' - ', round(exp(`97.5%`),2), ')'))
kable(summary.jm) %>% kable_styling()
```

# Joint Model 4 HR's
```{r jm4-hr}
summary.jm <- as.data.frame(summary(jointFit.comp.4)$Survival)
summary.jm <- summary.jm %>% mutate(HR = exp(Mean))
summary.jm <- round(summary.jm,3)
summary.jm <- summary.jm %>%mutate(P = ifelse(P == 0, '< 0.000', P))
summary.jm <- summary.jm %>% mutate(HR = paste0(round(exp(Mean), 2), '(', round(exp(`2.5%`),2), ' - ', round(exp(`97.5%`),2), ')'))
kable(summary.jm) %>% kable_styling()
```

# Joint Model 6 HR's
```{r jm6-hr}
summary.jm <- as.data.frame(summary(jointFit.comp.6)$Survival)
summary.jm <- summary.jm %>% mutate(HR = exp(Mean))
summary.jm <- round(summary.jm,3)
summary.jm <- summary.jm %>%mutate(P = ifelse(P == 0, '< 0.000', P))
summary.jm <- summary.jm %>% mutate(HR = paste0(round(exp(Mean), 2), '(', round(exp(`2.5%`),2), ' - ', round(exp(`97.5%`),2), ')'))
kable(summary.jm) %>% kable_styling()
```

#  ROC Joint Model 1 (Time Dependant) - ROC and tvAUC, Prediction at one year using data at month 0
```{r roc-1-12}
roc <- tvROC(jointFit.comp.1, newdata = ntprobnp_survival, Tstart = 0, Dt = 12)
roc
plot(roc)
tvAUC(roc)
```

#  ROC Joint Model 1 (Time Dependant) - ROC and tvAUC Prediction at month 24 (one year) using data up until month 12
```{r roc-1-24}
roc <- tvROC(jointFit.comp.1, newdata = ntprobnp_survival, Tstart = 12, Dt = 12)
roc
plot(roc)
tvAUC(roc)

```

#  ROC Joint Model 4 (Time Dependant + Time Dependant Slope) - ROC and tvAUC, Prediction at one year using data at month 0
```{r roc-4-12}
roc_4 <- tvROC(jointFit.comp.4, newdata = ntprobnp_survival, Tstart = 0, Dt = 12)
roc_4
plot(roc_4)
tvAUC(roc_4)
```


#  ROC Joint Model 4 (Time Dependant + Time Dependant Slope) - ROC and tvAUC Prediction at month 24 (one year) using data up until month 12
```{r roc-4-24}
roc_4 <- tvROC(jointFit.comp.4, newdata = ntprobnp_survival, Tstart = 12, Dt = 12)
roc_4
plot(roc_4)
tvAUC(roc_4)

```

#  ROC Joint Model 6 (Area) - ROC and tvAUC, Prediction at one year using data at month 0
```{r roc-6-12}
roc_6 <- tvROC(jointFit.comp.6, newdata = ntprobnp_survival, Tstart = 0, Dt = 12)
roc_6
plot(roc_6)
tvAUC(roc_6)
```

#  ROC Joint Model 6 (Area) - ROC and tvAUC Prediction at month 24 (one year) using data up until month 12
```{r roc-6-24}
roc_6 <- tvROC(jointFit.comp.6, newdata = ntprobnp_survival, Tstart = 12, Dt = 12)
roc_6
plot(roc_6)
tvAUC(roc_6)

```

# Calibration Plot for Joint Model 1 (Time Dependant) at one year using data at month 0
```{r calibration-plot-1-12}
calibration_plot(jointFit.comp.1, newdata = ntprobnp_survival, Tstart = 0, Dt = 12, main = "Calibration Curve for Joint Model 1 (Time Dependant) at 12 Months using Data at month 0")

```

# Calibration Plot for Joint Model 1 (Time Dependant) at month 24 (one year) using data up until month 12
```{r calibration-plot-1-24}
calibration_plot(jointFit.comp.1, newdata = ntprobnp_survival, Tstart = 12, Dt = 12, main = "Calibration Curve for Joint Model 1 (Time Dependant) at month 24 using Data at month 12")

```

# Calibration Plot for Joint Model 4 (Time Dependant + Time Dependant Slope) at one year using data at month 0
```{r calibration-plot-4-12}
calibration_plot(jointFit.comp.4, newdata = ntprobnp_survival, Tstart = 0, Dt = 12, main = "Calibration Curve for Joint Model 4 (Time Dependant + Time Dependant Slope) at 12 Months using Data at month 0")

```

# Calibration Plot for Joint Model 4 (Time Dependant + Time Dependant Slope) at month 24 (one year) using data up until month 12
```{r calibration-plot-4-24}
calibration_plot(jointFit.comp.4, newdata = ntprobnp_survival, Tstart = 12, Dt = 12, main = "Calibration Curve for Joint Model 4 (Time Dependant + Time Dependant Slope) at month 24 using Data at month 12")

```

# Calibration Plot for Joint Model 6 (Area) at one year using data at month 0
```{r calibration-plot-6-12}
calibration_plot(jointFit.comp.6, newdata = ntprobnp_survival, Tstart = 0, Dt = 12, main = "Calibration Curve for Joint Model 6 (Area) at 12 Months using Data at month 0")

```

# Calibration Plot forJoint Model 6 (Area) at month 24 (one year) using data up until month 12
```{r calibration-plot-6-24}
calibration_plot(jointFit.comp.6, newdata = ntprobnp_survival, Tstart = 12, Dt = 12, main = "Calibration Curve for Joint Model 6 (Area) at month 24 using Data at month 12")

```

# Brier Score for Joint Model 1 (Time Dependant) at one year using data at month 0
```{r brier-score-1-12}
pErr.1.12 <- tvBrier(jointFit.comp.1, newdata = ntprobnp_survival, Tstart = 0, Dt = 12)
pErr.1.12
```

# Brier Score for Joint Model 1 (Time Dependant) at month 24 (one year) using data up until month 12
```{r brier-score-1-24}
pErr.1.24 <- tvBrier(jointFit.comp.1, newdata = ntprobnp_survival, Tstart = 12, Dt = 12)
pErr.1.24
```

# Brier Score for Joint Model 4 (Time Dependant + Time Dependant Slope) at one year using data at month 0
```{r brier-score-4-12}
pErr.4.12 <- tvBrier(jointFit.comp.4, newdata = ntprobnp_survival, Tstart = 0, Dt = 12)
pErr.4.12
```

# Brier Score for Joint Model 4 (Time Dependant + Time Dependant Slope) at month 24 (one year) using data up until month 12
```{r brier-score-4-24}
pErr.4.24 <- tvBrier(jointFit.comp.4, newdata = ntprobnp_survival, Tstart = 12, Dt = 12)
pErr.4.24
```

# Brier Score for Joint Model 6 (Area) at one year using data at month 0
```{r brier-score-6-12}
pErr.6.12 <- tvBrier(jointFit.comp.6, newdata = ntprobnp_survival, Tstart = 0, Dt = 12)
pErr.6.12
```

# Brier Score for Joint Model 6 (Area) at month 24 (one year) using data up until month 12
```{r brier-score-6-24}
pErr.6.24 <- tvBrier(jointFit.comp.6, newdata = ntprobnp_survival, Tstart = 12, Dt = 12)
pErr.6.24
```

# Validation - Atmosphere Import
```{r atmospher import}

# Import Lab Results
lrb <- import("Atmosphere/STATA1/lrb.dta")
# Import Baseline data
baseline_atmosphere <- import("Atmosphere/Master/atmosphere master.dta")

# Filter Natriuretic Peptides
NatriureticPeptides <- lrb %>% filter(grepl("BNP", PARNAM1C))
lrb <- NULL

```

# Validation - Atmosphere Tidy
```{r atmospher-tidy}
# Filter NT-ProBNP
ntprobnp_atmosphere <- NatriureticPeptides %>% filter(PARNAM1C == "NTBNP")
# Select appropriate columns
ntprobnp_atmosphere <- ntprobnp_atmosphere %>% select(sid1a = SID1A,
                                ntprobnp_value = LABRSL1N,
                                sample_date = SMPCOL1O,
                                visit = VISNAM1A)

NatriureticPeptides <- NULL

baseline_atmosphere <- baseline_atmosphere %>% select(sid1a = sid1a,
                                randate =tstt_1o,
                                t2dth,
                                t2cnpt,
                                t2hhf,
                                c4dth,
                                c4cnpt,
                                c4hhf,
                                age = age_1n,
                                region,
                                sex = sex1c,
                                trt = trtr,
                                bmi = bmibl1n,
                                NYHA = nyhcls2c,
                                eGFR = egfrbl1n,
                                ejf = lvefsc1n,
                                diabetes = dbt3c,
                                atrial_fibrillation = af_flg1,
                                prior_hf = phh_flg1,
                                prior_mi = priormi,
                                prior_stroke = stroke_flg1,
                                ischemic_hf = isc_flg1,
                                sbp = sbpbl1n,
                                heart_rate = plsbl1n
)

md.pattern(baseline_atmosphere)

```

```{r atmosphere-data-cleaning}

# Recode for use with Paradigmn
baseline_atmosphere <- baseline_atmosphere %>% mutate(trt_drug = as.factor(case_when(
  trt == "A" ~ "LCZ",
  trt == "B" ~ "LCZ",
  trt == "C" ~ "Enalapril"
)))

# Recode Region to factor
baseline_atmosphere <- baseline_atmosphere %>% mutate(region = as.factor(case_when(
  region == 1 ~ "North American",
  region == 2 ~ "Latin America",
  region == 3 ~ "Western Europe",
  region == 4 ~ "Central Europe",
  region == 5 ~ "Pacific Asia/Pacific and Other"
)))

# Recode NYHA Class to factor
baseline_atmosphere <- baseline_atmosphere %>% mutate(NYHA_class = as.factor(case_when(
  NYHA == "1" ~ "CLASS_1",
  NYHA == "2" ~ "CLASS_2",
  NYHA == "3" ~ "CLASS_3",
  NYHA == "4" ~ "CLASS_4"
)))

# Recode Sex as Factor
baseline_atmosphere <- baseline_atmosphere %>% mutate(sex = as.factor(
  if_else(sex == 1, "MALE", "FEMALE")))

# Recode diabetes as Factor
baseline_atmosphere <- baseline_atmosphere %>% mutate(diabetes = as.factor(
  if_else(diabetes == 1, "TRUE", "FALSE")))

# Recode AF as Factor
baseline_atmosphere <- baseline_atmosphere %>% mutate(atrial_fibrillation = as.factor(
  if_else(atrial_fibrillation == 1, "TRUE", "FALSE")))

# Recode Prior Hospitalisation for Heart Failure as Factor
baseline_atmosphere <- baseline_atmosphere %>% mutate(prior_hf = as.factor(
  if_else(prior_hf == 1, "TRUE", "FALSE")))

# Recode Prior MI as Factor
baseline_atmosphere <- baseline_atmosphere %>% mutate(prior_mi = as.factor(
  if_else(prior_mi == 1, "TRUE", "FALSE")))

# Recode Prior Stroke as Factor
baseline_atmosphere <- baseline_atmosphere %>% mutate(prior_stroke = as.factor(
  if_else(prior_stroke == 1, "TRUE", "FALSE")))

# Recode Time to Composite Event to 28 Calendar Day Months
baseline_atmosphere<- baseline_atmosphere%>% mutate(composite_month = t2cnpt / 28)

# Invert Censor For Composite Event to be 1 for Event
baseline_atmosphere<- baseline_atmosphere%>% mutate(composite_status = if_else(c4cnpt == 1, 0, 1))

# Recode Ischemic HF to Factor
baseline_atmosphere <- baseline_atmosphere %>% mutate(ischemic_hf = as.factor(if_else(ischemic_hf == 1, "YES", "NO")))

# Remove Missing for Ischemic HF
baseline_atmosphere <- baseline_atmosphere %>% filter(!is.na(ischemic_hf))

# Join Baseline Characteristics with NT-ProBNP Values
ntprobnp_atmosphere <- ntprobnp_atmosphere %>% left_join(select(baseline_atmosphere, sid1a, trt_drug, eGFR, randate, age, sex, bmi, atrial_fibrillation, ischemic_hf), by="sid1a")

#################################################################
##                        Log ntprobnp                         ##
##  and filter any patients who did not undergo randomisation  ##
#################################################################
ntprobnp_atmosphere <- ntprobnp_atmosphere %>% mutate(logNTProBNP = log(ntprobnp_value)) %>%
  filter(!is.na(randate)) 

#######################################################################
##  Calculate Months Since Randomisation using 28 day calendar month  ##
#######################################################################
ntprobnp_atmosphere <- ntprobnp_atmosphere %>% 
  mutate(sample_date_formatted = as_date(sample_date, format = "%d%b%Y")) %>%
  mutate(sample_days = interval(randate, sample_date_formatted) %/% days(1)) %>% mutate(sample_month = sample_days / 28)

##################################################################
##        filter any observations prior to randomisation        ##
##################################################################
ntprobnp_atmosphere <- ntprobnp_atmosphere %>% filter(sample_month >= 0)

# Filter Missing NT-ProBNP values
ntprobnp_atmosphere <- ntprobnp_atmosphere %>% filter(!is.na(logNTProBNP))

# Arrange Data
ntprobnp_atmosphere <- ntprobnp_atmosphere %>% arrange(sid1a) %>% group_by(sid1a) %>% arrange(sample_days, .by_group = TRUE)

# Filter Patients Removed from NT-ProBNP Dataset from Baseline
baseline_atmosphere <- baseline_atmosphere %>% filter(sid1a %in% ntprobnp_atmosphere$sid1a)

# See Missing Values
missing_data <- ntprobnp_atmosphere[!complete.cases(ntprobnp_atmosphere),]

# Filter Patients with Missing NYHA Class, BMI or NT-ProBNP
baseline_atmosphere <- baseline_atmosphere %>% filter(!is.na(NYHA_class)) %>%
  filter(!is.na(bmi))

# Sort Two Datasets to same order
ntprobnp_atmosphere <- ntprobnp_atmosphere[order(ntprobnp_atmosphere$sid1a),]
row.names(ntprobnp_atmosphere) <- 1:nrow(ntprobnp_atmosphere)
baseline_atmosphere <- baseline_atmosphere[order(baseline_atmosphere$sid1a),]
row.names(baseline_atmosphere) <- 1:nrow(baseline_atmosphere)

# Filter Any Patients Removed from Either Dataset
ntprobnp_atmosphere <- ntprobnp_atmosphere %>% filter(sid1a %in% baseline_atmosphere$sid1a)
baseline_atmosphere <- baseline_atmosphere %>% filter(sid1a %in% ntprobnp_atmosphere$sid1a)

```

```{r atmosphere-merge}
# Merge all Baseline Characteristics to the NT-ProBNp Values
atmosphereSurvivalNTProBNP <- ntprobnp_atmosphere %>% left_join(baseline_atmosphere, by = c("sid1a", "trt_drug", "eGFR", "randate", "age", "sex", "bmi", "atrial_fibrillation", "ischemic_hf"))

export(atmosphereSurvivalNTProBNP, "r-scripts-ryan/Final-Analysis/ALL-NT-PROBNP/Data/atmospherSurvivalNTProBNP.rds")

```

# Validation ROC Joint Model 1 (Time Dependant) - ROC and tvAUC, Prediction at one year using data at month 0
```{r roc-atmosphere-1-12}
roc <- tvROC(jointFit.comp.1, newdata = atmosphereSurvivalNTProBNP, Tstart = 0, Dt = 12)
roc
plot(roc)
tvAUC(roc)
```

# Validation ROC Joint Model 1 (Time Dependant) - ROC and tvAUC Prediction at month 24 (one year) using data up until month 12
```{r roc-atmosphere-1-24}
roc <- tvROC(jointFit.comp.1, newdata = atmosphereSurvivalNTProBNP, Tstart = 12, Dt = 12)
roc
plot(roc)
tvAUC(roc)

```

# Validation ROC Joint Model 4 (Time Dependant + Time Dependant Slope) - ROC and tvAUC, Prediction at one year using data at month 0
```{r roc-atmosphere-4-12}
roc_4 <- tvROC(jointFit.comp.4, newdata = atmosphereSurvivalNTProBNP, Tstart = 0, Dt = 12)
roc_4
plot(roc_4)
tvAUC(roc_4)
```


# Validation ROC Joint Model 4 (Time Dependant + Time Dependant Slope) - ROC and tvAUC Prediction at month 24 (one year) using data up until month 12
```{r roc-atmosphere-4-24}
roc_4 <- tvROC(jointFit.comp.4, newdata = atmosphereSurvivalNTProBNP, Tstart = 12, Dt = 12)
roc_4
plot(roc_4)
tvAUC(roc_4)

```

# Validation ROC Joint Model 6 (Area) - ROC and tvAUC, Prediction at one year using data at month 0
```{r roc-atmosphere-6-12}
roc_6 <- tvROC(jointFit.comp.6, newdata = atmosphereSurvivalNTProBNP, Tstart = 0, Dt = 12)
roc_6
plot(roc_6)
tvAUC(roc_6)
```

# Validation ROC Joint Model 6 (Area) - ROC and tvAUC Prediction at month 24 (one year) using data up until month 12
```{r roc-atmosphere-6-24}
roc_6 <- tvROC(jointFit.comp.6, newdata = atmosphereSurvivalNTProBNP, Tstart = 12, Dt = 12)
roc_6
plot(roc_6)
tvAUC(roc_6)

```

# Atmosphere Validation Calibration Plot for Joint Model 1 (Time Dependant) at one year using data at month 0
```{r validation-calibration-plot-1-12}
calibration_plot(jointFit.comp.1, newdata = atmosphereSurvivalNTProBNP, Tstart = 0, Dt = 12, main = "Calibration Curve for Joint Model 1 (Time Dependant) at 12 Months using Data at month 0")

```

# Atmosphere Validation Calibration Plot for Joint Model 1 (Time Dependant) at month 24 (one year) using data up until month 12
```{r validation-calibration-plot-1-24}
calibration_plot(jointFit.comp.1, newdata = atmosphereSurvivalNTProBNP, Tstart = 12, Dt = 12, main = "Calibration Curve for Joint Model 1 (Time Dependant) at month 24 using Data at month 12")

```

# Atmosphere Validation Calibration Plot for Joint Model 4 (Time Dependant + Time Dependant Slope) at one year using data at month 0
```{r validation-calibration-plot-4-12}
calibration_plot(jointFit.comp.4, newdata = atmosphereSurvivalNTProBNP, Tstart = 0, Dt = 12, main = "Calibration Curve for Joint Model 4 (Time Dependant + Time Dependant Slope) at 12 Months using Data at month 0")

```

# Atmosphere Validation Calibration Plot for Joint Model 4 (Time Dependant + Time Dependant Slope) at month 24 (one year) using data up until month 12
```{r validation-calibration-plot-4-24}
calibration_plot(jointFit.comp.4, newdata = atmosphereSurvivalNTProBNP, Tstart = 12, Dt = 12, main = "Calibration Curve for Joint Model 4 (Time Dependant + Time Dependant Slope) at month 24 using Data at month 12")

```

# Atmosphere Validation Calibration Plot for Joint Model 6 (Area) at one year using data at month 0
```{r validation-calibration-plot-6-12}
calibration_plot(jointFit.comp.6, newdata = atmosphereSurvivalNTProBNP, Tstart = 0, Dt = 12, main = "Calibration Curve for Joint Model 6 (Area) at 12 Months using Data at month 0")

```

# Atmosphere Validation Calibration Plot forJoint Model 6 (Area) at month 24 (one year) using data up until month 12
```{r validation-calibration-plot-6-24}
calibration_plot(jointFit.comp.6, newdata = atmosphereSurvivalNTProBNP, Tstart = 12, Dt = 12, main = "Calibration Curve for Joint Model 6 (Area) at month 24 using Data at month 12")

```

# Atmosphere Validation Brier Score for Joint Model 1 (Time Dependant) at one year using data at month 0
```{r validation-brier-score-1-12}
pErr.1.12 <- tvBrier(jointFit.comp.1, newdata = atmosphereSurvivalNTProBNP, Tstart = 0, Dt = 12)
pErr.1.12
```

# Atmosphere Validation Brier Score for Joint Model 1 (Time Dependant) at month 24 (one year) using data up until month 12
```{r validation-brier-score-1-24}
pErr.1.24 <- tvBrier(jointFit.comp.1, newdata = atmosphereSurvivalNTProBNP, Tstart = 12, Dt = 12)
pErr.1.24
```

# Atmosphere Validation Brier Score for Joint Model 4 (Time Dependant + Time Dependant Slope) at one year using data at month 0
```{r validation-brier-score-4-12}
pErr.4.12 <- tvBrier(jointFit.comp.4, newdata = atmosphereSurvivalNTProBNP, Tstart = 0, Dt = 12)
pErr.4.12
```

# Atmosphere Validation Brier Score for Joint Model 4 (Time Dependant + Time Dependant Slope) at month 24 (one year) using data up until month 12
```{r validation-brier-score-4-24}
pErr.4.24 <- tvBrier(jointFit.comp.4, newdata = atmosphereSurvivalNTProBNP, Tstart = 12, Dt = 12)
pErr.4.24
```

# Atmosphere Validation Brier Score for Joint Model 6 (Area) at one year using data at month 0
```{r validation-brier-score-6-12}
pErr.6.12 <- tvBrier(jointFit.comp.6, newdata = atmosphereSurvivalNTProBNP, Tstart = 0, Dt = 12)
pErr.6.12
```

# Atmosphere Validation Brier Score for Joint Model 6 (Area) at month 24 (one year) using data up until month 12
```{r validation-brier-score-6-24}
pErr.6.24 <- tvBrier(jointFit.comp.6, newdata = atmosphereSurvivalNTProBNP, Tstart = 12, Dt = 12)
pErr.6.24
```

# Clear the joint models from memory
```{r clear-memory}
jointFit.comp.1 <- NULL
jointFit.comp.4 <- NULL
jointFit.comp.6 <- NULL
```

# Fit Last Measurement Cox PH Model
```{r comparative-analysis-cox-last}
ntprobnp.last <- ntprobnp %>% group_by(sid1a) %>% arrange(sid1a, sample_month) %>%
  mutate(row_number = row_number()) %>% filter(row_number == n())

ntprobnp.last <- ntprobnp.last %>% left_join(baseline, by = c(
  "sid1a",
  "trt_drug",
  "age",
  "atrial_fibrillation",
  "bmi",
  "sex",
  "eGFR",
  "randate",
  "ischemic_hf"
))

coxPH.comp.fit.last <- coxph(Surv(composite_month, composite_status) ~
                     trt_drug +
                     age +
                     sex +
                     region +
                     bmi + 
                     eGFR + 
                     ejf +
                     NYHA_class +
                     diabetes + 
                     sbp +
                     heart_rate +
                     atrial_fibrillation +
                     ischemic_hf +
                     prior_hf +
                     prior_mi + 
                     prior_stroke +
                     logNTProBNP,
                   data = ntprobnp.last,
                   x = TRUE)

summary(coxPH.comp.fit.last)

```

# Last Measurement Cox PH Summary
```{r summary-cox-ph-last}

cox.last.summary <- as.data.frame(summary(coxPH.comp.fit.last)$conf.int) %>%
  rownames_to_column() %>%
  mutate(
    `Pr(>|z|)` = as.data.frame(summary(coxPH.comp.fit.last)$coefficients)$`Pr(>|z|)`
  ) %>%
  mutate(
    `exp(coef)` = round(`exp(coef)`, 2),
    `lower .95` = round(`lower .95`, 2),
    `upper .95` = round(`upper .95`, 2),
  `p-value` = if_else(round(`Pr(>|z|)`,3) == 0, "<0.001", as.character(round(`Pr(>|z|)`,3)))) %>%
  dplyr::select(rowname,
  HR = `exp(coef)`,
  `lower .95`,
  `upper .95`,
  `p-value`
  ) %>% mutate(
    HR = paste0(HR, " (", `lower .95`, "-", `upper .95`, ")")
  )

kable(cox.last.summary)

```

# Cox PH Last AUC 12 Month
```{r cox-ph-last-auc-12}
library(risksetROC)

ntprobnp.last.na.rm <- ntprobnp.last
ntprobnp.last.na.rm$n <- 1:nrow(ntprobnp.last.na.rm)
ntprobnp.last.na.rm <- ntprobnp.last.na.rm %>% filter(!n %in% coxPH.comp.fit.last$na.action)

surivival.time = ntprobnp.last.na.rm$composite_month
status = ntprobnp.last.na.rm$composite_status
eta = coxPH.comp.fit.last$linear.predictors
method = "Cox"

last.auc.12 <- risksetROC(Stime = surivival.time, status = status, method = method, predict.time = 12, marker = eta, weight = "conditional", xlab = '1 - Specificity', ylab = "Sensitivity")

last.auc.12$AUC

kable(list(AUC = last.auc.12$AUC), col.names = c("AUC"))

```

# Cox PH Last AUC 24 Months
```{r cox-ph-last-auc-24}

last.auc.24 <- risksetROC(Stime = surivival.time, status = status, method = method, predict.time = 24, marker = eta, weight = "conditional", xlab = '1 - Specificity', ylab = "Sensitivity")

last.auc.24$AUC

kable(list(AUC = last.auc.24$AUC), col.names = c("AUC"))

```

# Fit extended Cox PH
```{r, comparative-analysis-extended-cox}

ntprobnp.extended <- ntprobnp %>% group_by(sid1a) %>% left_join(baseline, by = c(
  "sid1a",
  "trt_drug",
  "age",
  "atrial_fibrillation",
  "bmi",
  "sex",
  "eGFR",
  "randate",
  "ischemic_hf"
)) %>% arrange(sid1a, sample_month) %>%
  mutate(
    comp_start = sample_month,
    comp_stop = lead(sample_month, default = last(composite_month)),
    comp_event = if_else(sample_month == last(sample_month), composite_status, 0)
  )

extendedCoxPH.comp.fit <- coxph(Surv(comp_start, comp_stop, comp_event) ~
                     trt_drug +
                     age +
                     sex +
                     region +
                     bmi + 
                     eGFR + 
                     ejf +
                     NYHA_class +
                     diabetes + 
                     sbp +
                     heart_rate +
                     atrial_fibrillation +
                     ischemic_hf +
                     prior_hf +
                     prior_mi + 
                     prior_stroke +
                     logNTProBNP,
                   data = ntprobnp.extended,
                   x = TRUE)

summary(extendedCoxPH.comp.fit)

```

# Extented Cox PH Summary
```{r summary-cox-ph-extended}

cox.extended.summary <- as.data.frame(summary(extendedCoxPH.comp.fit)$conf.int) %>%
  rownames_to_column() %>%
  mutate(
    `Pr(>|z|)` = as.data.frame(summary(extendedCoxPH.comp.fit)$coefficients)$`Pr(>|z|)`
  ) %>%
  mutate(
    `exp(coef)` = round(`exp(coef)`, 2),
    `lower .95` = round(`lower .95`, 2),
    `upper .95` = round(`upper .95`, 2),
  `p-value` = if_else(round(`Pr(>|z|)`,3) == 0, "<0.001", as.character(round(`Pr(>|z|)`,3)))) %>%
  dplyr::select(rowname,
  HR = `exp(coef)`,
  `lower .95`,
  `upper .95`,
  `p-value`
  ) %>% mutate(
    HR = paste0(HR, " (", `lower .95`, "-", `upper .95`, ")")
  )

kable(cox.extended.summary)

```

# AUC Extended Cox PH, Prediction at one year
```{r extended-auc-12}

ntprobnp.extended.na.rm <- ntprobnp.extended
ntprobnp.extended.na.rm$n <- 1:nrow(ntprobnp.extended.na.rm)
ntprobnp.extended.na.rm <- ntprobnp.extended.na.rm %>% filter(!n %in% extendedCoxPH.comp.fit$na.action)

surivival.time = ntprobnp.extended.na.rm$comp_stop
entry.time = ntprobnp.extended.na.rm$comp_start
status = ntprobnp.extended.na.rm$comp_event
eta = extendedCoxPH.comp.fit$linear.predictors
method = "Cox"

extended.auc.12 <- risksetROC(Stime = surivival.time, entry = entry.time, status = status, method = method, predict.time = 12, marker = eta, weight = "conditional", xlab = '1 - Specificity', ylab = "Sensitivity")

extended.auc.12$AUC
kable(list(AUC = extended.auc.12$AUC), col.names = c("AUC"))
```

# AUC Extended Cox PH Prediction at 24 months
```{r extended-auc-24}
extended.auc.24 <- risksetROC(Stime = surivival.time, entry = entry.time, status = status, method = method, predict.time = 24, marker = eta, weight = "conditional", xlab = '1 - Specificity', ylab = "Sensitivity")

extended.auc.24$AUC
kable(list(AUC = extended.auc.24$AUC), col.names = c("AUC"))
```
